Reproducible Research - Course Project 1
----------------------------------------

This is my submission for this course project. You can see the rest of the submission here. (https://github.com/BoKaplan/RepData_PeerAssessment1)
  
<br>

###Loading and preprocessing the data
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Import data
act <- read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))
```
```{r, echo=TRUE}
# Some variables
str(act)
```


The variables included in this dataset are:  
1. **Steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)  
2. **Date**: The date on which the measurement was taken in YYYY-MM-DD format  
3. **Interval**: Identifier for the 5-minute interval in which measurement was taken

<br>

###What is mean total number of steps taken per day?
**1. Calculate the total number of steps taken per day**

```{r echo=TRUE}
# Create and print number of steps per day
StepsPerDay <- aggregate(act$steps, list(act$date), FUN=sum)
colnames(StepsPerDay) <- c("Date", "Steps")
StepsPerDay
```


**2. Create histogram**

```{r histogram1, echo=TRUE, fig.width=10, warning=FALSE}
# Drawing the histogram
sum_by_date <- tapply(act$steps,act$date,sum,na.rm=TRUE)
hist(sum_by_date,col=heat.colors(8),xlab="Number of Steps",main="Histogram of Total Steps Per Day")
```


**3. Figure out what the mean and median are**

```{r echo=TRUE}
# Mean
mean(StepsPerDay$Steps, na.rm = TRUE)
# Median
median(StepsPerDay$Steps, na.rm = TRUE)
```

<br>

###What is the average daily activity pattern?
**1. Create a time series plot**

```{r timeplot1, echo=TRUE, fig.width=10, warning=FALSE}
# Create a table
StepsPerInterval <- tapply(act$steps,act$interval,mean,na.rm=TRUE)
# Create a line plot
plot(row.names(StepsPerInterval),StepsPerInterval,type="l",xlab="Time Intervals (in Minutes)",ylab="Average Number of Steps",main="Time Series Plot of Average Daily Steps")
```

**2. Maximum number of steps with 5-minute intervals**

```{r echo=TRUE, fig.width=10, warning=FALSE}
# Interval
maxInterval <- names(sort(StepsPerInterval, decreasing = TRUE)[1])
# Steps
maxSteps <- sort(StepsPerInterval, decreasing = TRUE)[1]
```

```{r echo=TRUE}
# Missing values
x <- max(StepsPerInterval)
match(x,StepsPerInterval)
```

```{r echo=TRUE}
StepsPerInterval[104]
```

<br>

###Imputting missing values
**1. Figure out the missing values**
```{r echo=TRUE}
# Missing values
sum(is.na(act))
```

**2. Adjust dataset**

```{r echo=TRUE}
# Create new dataset
act_na <- act[is.na(act),]
act_no_na <- act[complete.cases(act),]
act_na$steps <- as.numeric(StepsPerInterval)
new_act <- rbind(act_na,act_no_na)
new_act <- new_act[order(new_act[,2],new_act[,3]),]
```

**3. Adjust steps per day**

```{r echo=TRUE}
# Adjust steps
AdjustedStepsPerDay <- aggregate(new_act$steps, list(new_act$date), FUN=sum)
colnames(AdjustedStepsPerDay) <- c("Date", "Steps")
AdjustedStepsPerDay
```

**4. Create adjusted histogram**

```{r histogram2, echo=TRUE, fig.width=10, warning=FALSE}
# Drawing the histogram
new_sum_by_date <- tapply(new_act$steps,new_act$date,sum)
hist(new_sum_by_date,col=terrain.colors(8),xlab="Total Steps by Date",main="Adjusted Histogram of Total Steps by Date (No Missing Values)")
```

**5. Figure out the adjusted mean and median**

```{r echo=TRUE}
# Mean
mean(AdjustedStepsPerDay$Steps, na.rm = TRUE)
# Median
median(AdjustedStepsPerDay$Steps, na.rm = TRUE)
```

<br>

###Are there differences in activity patterns between weekdays and weekends?
**1. Create a new factor variable in the dataset with two levels**

```{r echo=TRUE}
# Creating a new dataset
is_weekday <- function(d) {
  wd <- weekdays(d)
  ifelse (wd == "Saturday" | wd == "Sunday", "Weekend", "Weekday")
}

wx <- sapply(new_act$date, is_weekday)
new_act$wk <- as.factor(wx)
head(new_act)
```
**2. Create two time series plot for the average number of steps during weekdays and weekends**

```{r timeplot2, echo=TRUE, fig.width=10, warning=FALSE}
# Create table
wdwx_act <- aggregate(steps ~ wk+interval, data=new_act, FUN=mean)
# Create a line plot
library(lattice)
xyplot(steps ~ interval | factor(wk),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of Steps",
       main="Activity Pattern (Weekdays vs Weekends)",
       type="l",
       lty=1,
       data=wdwx_act)
```
